# Module 1

<!-- ====== COLOURED BANNER ====== -->
```{r, results='asis', echo=FALSE}
banner_box(
  "Module Focus: Introducing the foundational ideas",
  "The focus is on understanding what data are, how they are collected, and how they should be summarised and interpreted responsibly."
)
```
<!-- ====== END BANNER ====== -->

## Variables, Types of Data, and Sampling

Understanding data begins with understanding variables.

**Statistical terminology**

- **Cases**: the individuals or objects being studied
- **Variables**: characteristics measured on each case
- **Data**: the observed values of the variables

**Types of data**

| Data type | Description | Examples |
|---------|------------|----------|
| Quantitative | Numerical values with meaningful units | Height (cm), pulse rate |
| Categorical (Nominal) | Categories with no natural order | Gender, faculty |
| Ordinal | Categories with a natural order | Likert scale responses |
| Identifier | Unique labels for each case | Student number |

Categorical variables are often **coded numerically** in software, but the numbers themselves have no numerical meaning.

**Populations and Sampling**
Because studying an entire population is usually impractical, statistics relies on samples. Valid conclusions depend on three key ideas:

1. **Sampling** – samples should be representative of the population.

2. **Randomising** – random selection reduces bias and allows generalisation.

3. **Sample size** – larger samples tend to give more accurate results, but at higher cost.

Only **random samples** support valid statistical inference.

---

## Visualising Data: One Categorical Variable

The **distribution of a variable** describes what values it takes and how often they occur. Graphs allow large amounts of data to be summarised clearly and patterns to be identified.

**Numerical summaries**

- Frequency tables for categorical variables
- Use pivot tables in Excel

**Bar charts**

- Used for **categorical variables**
- Bars do **not** touch
- Order bars logically (alphabetical or by size)

**Pie charts**

- Not recommended as difficult for the reader to fully understand
- Show **parts of a whole**
- Must include all categories (total = 100%)
- Best used sparingly and with few categories

---

## Contingency Tables 

When analysing **two categorical variables**, we use a **contingency table** (also called a two-way table or cross-tabulation).

**Types of distributions**

- **Joint distribution**: proportions of all cases in each cell
- **Marginal distribution**: proportions for one variable only (row or column totals)
- **Conditional distribution**: proportions within a subgroup

**Assessing association**

Two categorical variables are **associated** if the conditional distributions differ between groups. If the conditional distributions are similar, the variables are considered **independent**.

**Graphical support**

- **Stacked bar charts** are useful for comparing conditional distributions visually.


```{r, results='asis', echo=FALSE}
tip_box(
  "To assess association:",
  "Always compare **conditional distributions**, not just raw counts."
)
```
---

## Use, Abuse, and Misuse of Statistics

Statistics are powerful, but they can be misleading if used poorly or dishonestly.

**Key ideas**

- Results can be **misinterpreted** through poor sampling or misleading summaries.
- Graphs can strongly influence interpretation depending on:
  - scale and axis limits
  - graph type
  - use of 3D effects or distorted perspectives

**Good data visualisation principles**

- Match the graph type to the **data type**
- Label axes clearly (variable names and units)
- Use honest scales (usually starting at zero)
- Avoid unnecessary decoration (especially 3D graphs)
- Aim to **tell the truth** and communicate clearly


A graph is only as good as the data it displays. No amount of visual design can fix poor or biased data.
---

# Review Questions: Module 1 {-}
```{r, results='asis', echo=FALSE}
banner_box(
  "Review Questions",
  "Once you have watched the lecture recordings you should be able to answer the following *Threshold* questions (solutions can be found at the end of this Workbook)."
)
```


<!-- HTML VERSION -->
```{=html}

<style>
  .revq { max-width:900px; margin:0 auto; }
  .revq .muted { color:#666; font-size:.9rem; }
  .revq-row { display:grid; grid-template-columns: 1fr 120px; gap:10px; align-items:start; margin:.6rem 0; }
  .revq textarea { width:100%; min-height:3.2rem; padding:6px; box-sizing:border-box; }
  .revq .chk { display:flex; align-items:center; gap:.5rem; }
  .revq .btn { padding:6px 10px; border:1px solid #ccc; border-radius:4px; background:#f7f7f7; cursor:pointer; }
  .revq .btn + .btn { margin-left:.4rem; }
  .revq-answers { background:#f6fbff; border-left:6px solid #2b8cbe; padding:.9rem 1rem; border-radius:6px; margin-top:.8rem; display:none; }
  .revq-answers h5 { margin:.2rem 0 .5rem; font-size:1.05rem; }
  .revq-answers p { margin:.4rem 0; }
</style>

<div class="revq" id="revq">
  <p class="muted">
    Write your answers in the boxes below. Tick the checkbox if you believe your answer is correct.
    Click <em>Show model answers</em> to compare.
  </p>

  (i) What is a variable? What is a value of a variable? Give examples.
  <div class="revq-row">
    <textarea id="revq_m1_q1" placeholder="Your answer…"></textarea>
    <label class="chk"><input type="checkbox"> Marked correct</label>
  </div>

  (ii) How can we tell if a variable is quantitative, categorical, ordinal, or an identifier?
  <div class="revq-row">
    <textarea id="revq_m1_q2" placeholder="Your answer…"></textarea>
    <label class="chk"><input type="checkbox"> Marked correct</label>
  </div>

  (iii) What is the difference between cross-sectional and time series data?
  <div class="revq-row">
    <textarea id="revq_m1_q3" placeholder="Your answer…"></textarea>
    <label class="chk"><input type="checkbox"> Marked correct</label>
  </div>

  (iv) What is the difference between a representative sample and a convenience sample? Advantages/disadvantages?
  <div class="revq-row">
    <textarea id="revq_m1_q4" placeholder="Your answer…"></textarea>
    <label class="chk"><input type="checkbox"> Marked correct</label>
  </div>

  (v) What is a contingency (two-way) table? When is it used?
  <div class="revq-row">
    <textarea id="revq_m1_q5" placeholder="Your answer…"></textarea>
    <label class="chk"><input type="checkbox"> Marked correct</label>
  </div>

  (vi) What is meant by a marginal, joint, and conditional distribution?
  <div class="revq-row">
    <textarea id="revq_m1_q6" placeholder="Your answer…"></textarea>
    <label class="chk"><input type="checkbox"> Marked correct</label>
  </div>

  (vii) List three things that may result in a bad graph.
  <div class="revq-row">
    <textarea id="revq_m1_q7" placeholder="Your answer…"></textarea>
    <label class="chk"><input type="checkbox"> Marked correct</label>
  </div>

  <div style="margin-top:.6rem;">
    <button class="btn" id="revq-show">Show model answers</button>
    <button class="btn" id="revq-hide" style="display:none;">Hide model answers</button>
    <button class="btn" id="revq-reset">Reset</button>
  </div>

  <div class="revq-answers" id="revq-model">
    <h5>Model answers</h5>

    <p><strong>(i)</strong> A <em>variable</em> is any characteristic of a case (an individual or object being described). A <em>value</em> of a variable is the observed measurement or category for that case. For example, for the variable <em>height</em>, values are measurements in cm or m. For <em>country of origin</em>, values may be Australia or Not Australia.</p>

    <p><strong>(ii)</strong> Quantitative variables take on numerical values for which mathematical operations (like +, ÷) make sense, with a unit of measurement. Categorical variables define categories. Ordinal variables are categorical with a natural order. Identifiers uniquely label each case.</p>

    <p><strong>(iii)</strong> Cross-sectional data measures values of multiple variables at a single point in time. Time series data measures one quantitative variable across regular time intervals.</p>

    <p><strong>(iv)</strong> Representative sample: random sample that reflects the population. Advantage = unbiased; Disadvantage = may be complicated to obtain. Convenience sample: includes individuals convenient for the researcher. Advantage = easy; Disadvantage = biased.</p>

    <p><strong>(v)</strong> A contingency (two-way) table visually shows the distribution of individuals across two categorical variables.</p>

    <p><strong>(vi)</strong> Marginal distribution = proportion in a row/column. Joint distribution = proportion in a cell. Conditional distribution = proportion in a row/column for a specific cell.</p>

    <p><strong>(vii)</strong> Possible reasons for a bad graph include: misleading scales, missing labels, too much clutter, inappropriate chart type, or truncated axes. See lecture notes for full details.</p>
  </div>
</div>

<script>
(function(){
  const show = document.getElementById("revq-show");
  const hide = document.getElementById("revq-hide");
  const model = document.getElementById("revq-model");
  const reset = document.getElementById("revq-reset");

  show.onclick = () => { model.style.display="block"; show.style.display="none"; hide.style.display="inline-block"; };
  hide.onclick = () => { model.style.display="none"; hide.style.display="none"; show.style.display="inline-block"; };
  reset.onclick = () => {
    document.querySelectorAll("#revq textarea").forEach(t => t.value="");
    document.querySelectorAll("#revq input[type=checkbox]").forEach(c => c.checked=false);
    model.style.display="none"; hide.style.display="none"; show.style.display="inline-block";
  };
})();
</script>
```


<!-- PDF Version -->
```{r, results='asis', echo=FALSE}
if (knitr::is_latex_output()) {
  cat('
\\textit{Write your answers in the spaces provided.}

\\medskip

\\textbf{(i) What is a variable? What is a value of a variable? Give examples of reach.}
\\begin{answerbox}\\vspace{3cm}\\end{answerbox}

\\textbf{(ii) How can we tell if a variable is quantitative, categorical, ordinal, or an identifier?}
\\begin{answerbox}\\vspace{3cm}\\end{answerbox}

\\textbf{(iii) What is the difference between cross sectional data and time series data?}
\\begin{answerbox}\\vspace{3cm}\\end{answerbox}

\\textbf{(iv) What is the difference between a representative sample and a convenience sample? How are they obtained and what are their relative advantages/disadvantages?}
\\begin{answerbox}\\vspace{3cm}\\end{answerbox}

\\textbf{(v) What is a contingency or two-way table? When is it used?}
\\begin{answerbox}\\vspace{3cm}\\end{answerbox}

\\textbf{(v) What is meant by a marginal distribution, a joint distribution and a conditional distribution?}
\\begin{answerbox}\\vspace{3cm}\\end{answerbox}

\\textbf{(v) List three things that may result in a bad graph.}
\\begin{answerbox}\\vspace{3cm}\\end{answerbox}
')
}
```
<!--
NOTE FOR FUTURE ME:
This module uses child Rmd files under section headings (##) so that:
- Module 1, Tutorial are in different files but render together
- Summary, Review, and Tutorial are numbered 1.1, 1.2, 1.3
- bs4_book splits these sections into separate HTML pages (via split_by: section)
- PDF output still renders cleanly with correct numbering

DO NOT add this file's children to _bookdown.yml
DO NOT use # headings in the child files
-->


```{r, child="01-Tut1.Rmd"}
```
