# Module 2

<!-- ====== COLOURED BANNER ====== -->
```{r, results='asis', echo=FALSE}
banner_box(
  "Module Focus: Exploring and understanding data",
  "The focus of this module is displaying one quantitative variable both visually and numberically."
)
```
<!-- ====== END BANNER ====== -->


## Gathering Data

**Types of Studies**
  
There are two main types of studies:
  
1. *Observational studies*
- Researchers observe and measure variables
- No intervention is applied
- Examples: surveys, polls, transactional data
- Can identify relationships but **cannot establish causation**
  
2. *Experimental studies*
- Researchers impose treatments and observe outcomes
- Can establish **cause and effect**
- *Explanatory variable (factor)*: the variable being manipulated  
- *Levels*: the values of the factor  
- *Treatments*: combinations of factor levels  
- *Experimental units*: individuals receiving treatments  
  

```{r, results='asis', echo=FALSE}
tip_box(
  "Only experiments can demonstrate causation.",
  "Observational studies can only show if variables are related."
)
```  

**Association vs Causation**
  
- *Association*: Two variables move together  
- *Causation*: One variable directly affects another  

A key issue is the presence of **lurking variables**:
  
- A **lurking variable** affects both variables but is not measured
- Can create misleading associations

**Example**  
Pet ownership and spending on cleaning products may be related, but a lurking variable (e.g. family with children) may explain the relationship.

**Types of Observational Studies**
  
- *Prospective study*: follows individuals into the future  
- *Retrospective study*: uses past/historical data 


**Principles of Good Experimental Design**
  
1. *Randomisation*
- Randomly assign subjects to treatments
- Reduces bias and lurking variables

2. *Control*
- Use control groups and placebos
- Allows fair comparison

3. *Replication*
- Use many subjects or repeat the experiment
- Reduces variability due to chance

4. *Blocking*
- Group similar individuals before assigning treatments
- Improves precision of comparisons

## Displaying Distributions
  
The **distribution of a quantitative variable** describes what values the data take
and how often those values occur.

**Key ideas**
  
- Raw data are often difficult to interpret without summarisation
- Graphical displays help reveal important patterns in the data
- Different graphs are suited to different data sizes and purposes

**Common displays for one quantitative variable**
  
*Histograms*  
- Best for larger data sets  
- Show the overall shape clearly  
- Require careful choice of bin width

*Boxplots*
- Summarise the distribution using the five-number summary  
- Excellent for comparing distributions across groups  
- Do not show detailed distribution shape

```{r, results='asis', echo=FALSE}
tip_box(
  "When constructing graphs",
  "It is important to include:
  
  - clear titles
  
  - labelled axes with units
  
  - appropriate scales
  
A graph should make the **overall features of the distribution easy to see**."
)
```

**Describing Distributions (Graphically)**
  
When interpreting a histogram or boxplot, describe:
  
1. **Shape**
- Symmetric, left-skewed, right-skewed  
- Number of peaks (modes)

2. **Centre**
- Typical value

3. **Spread**
- Range of values (exclude outliers if there are any)

4. **Outliers**
- Unusual observations

**Time Series Data**
  
- Data measured over time at regular intervals  
- Displayed using a **time series plot**
  
Use time series plots to:
- Identify trends
- Detect changes over time


## Describing Distributions with Numbers
  
**Summary Statistics**
  
We describe quantitative data using:
  
**Centre**

- Mean - Average value, sensitive to outliers

- Median - middle number when dataset ordered, resistant to outliers

**Spread**

- Range = max − min

- Interquartile Range (IQR) = [Q3 − Q1] - Used with the median for skewed data or data with outliers

- Standard deviation - Measures average spread around the mean, used with the mean for symmetric data


```{r, results='asis', echo=FALSE}
tip_box(
  "💡 Use:",
" - Mean/SD: for **symmetric distributions** with no outliers.
- Median/IQR: for **skewed distributions** or distributions with outilers."
)
```

**Quartiles and Five-Number Summary**
  
Five-number summary:

- Minimum  
- First quartile (Q1)  
- Median  
- Third quartile (Q3)  
- Maximum  

Used to summarise the entire distribution. These numbers for the boxplot.


**Choosing the Right Summary**
  
  | Distribution Type | Centre | Spread |
  |-----------------|--------|--------|
  | Symmetric        | Mean   | Standard deviation |
  | Skewed / Outliers| Median | IQR |
  
## Key Takeaways
  
- Observational studies show relationships, not causation  
- Experiments are required for cause-and-effect conclusions  
- Graphs help visualise distributions  
- Always describe distributions using: Shape, Centre, Spread, Outliers  
- Use appropriate numerical summaries based on distribution shape  

  
# Review Questions: Module 2 {-}

```{r, results='asis', echo=FALSE}
banner_box(
  "Review Questions",
  "Once you have watched the lecture recordings you should be able to answer the following *Threshold* questions (solutions can be found at the end of this Workbook)."
)
```

<!-- HTML VERSION -->
```{=html}

<style>
  .revq { max-width:900px; margin:0 auto; }
  .revq .muted { color:#666; font-size:.9rem; }
  .revq-row { display:grid; grid-template-columns: 1fr 120px; gap:10px; align-items:start; margin:.6rem 0; }
  .revq textarea { width:100%; min-height:3.2rem; padding:6px; box-sizing:border-box; }
  .revq .chk { display:flex; align-items:center; gap:.5rem; }
  .revq .btn { padding:6px 10px; border:1px solid #ccc; border-radius:4px; background:#f7f7f7; cursor:pointer; }
  .revq .btn + .btn { margin-left:.4rem; }
  .revq-answers { background:#f6fbff; border-left:6px solid #2b8cbe; padding:.9rem 1rem; border-radius:6px; margin-top:.8rem; display:none; }
</style>

<div class="revq" id="revq">

<p class="muted">
Write your answers in the boxes below. Tick the checkbox if you believe your answer is correct.
Click <em>Show model answers</em> to compare.
</p>

(i) What are the main differences between an observational study and an experiment?
<div class="revq-row">
  <textarea></textarea>
  <label class="chk"><input type="checkbox"> Marked correct</label>
</div>

(ii) State and define, with examples, the four principles of experimental design.
<div class="revq-row">
  <textarea></textarea>
  <label class="chk"><input type="checkbox"> Marked correct</label>
</div>

(iii) Which three types of graphs can be used to display the distribution of one quantitative variable?
<div class="revq-row">
  <textarea></textarea>
  <label class="chk"><input type="checkbox"> Marked correct</label>
</div>

(iv) What is an outlier?
<div class="revq-row">
  <textarea></textarea>
  <label class="chk"><input type="checkbox"> Marked correct</label>
</div>

(v) If the units of measurement are metres, what are the units of:
mean, median, standard deviation, and variance?
<div class="revq-row">
  <textarea></textarea>
  <label class="chk"><input type="checkbox"> Marked correct</label>
</div>

(vi) Which descriptions of centre and spread are best for a skewed distribution?
Which measures are best for a reasonably symmetric distribution?
<div class="revq-row">
  <textarea></textarea>
  <label class="chk"><input type="checkbox"> Marked correct</label>
</div>

<div style="margin-top:.6rem;">
  <button class="btn" id="revq-show">Show model answers</button>
  <button class="btn" id="revq-hide" style="display:none;">Hide model answers</button>
  <button class="btn" id="revq-reset">Reset</button>
</div>

<div class="revq-answers" id="revq-model">

<h5>Model answers</h5>

<p><strong>(i)</strong> Observational studies involve observing and measuring variables without intervention, while experiments impose treatments on individuals to observe effects. Observational studies can identify associations but cannot establish causation, whereas experiments can establish cause-and-effect relationships.</p>

<p><strong>(ii)</strong> The four principles are:
<br><strong>Randomisation:</strong> randomly assign subjects to treatments (reduces bias).
<br><strong>Control:</strong> use control groups/placebos to compare results.
<br><strong>Replication:</strong> use many subjects to reduce variability.
<br><strong>Blocking:</strong> group similar individuals before assigning treatments.
Example: In a drug trial, patients are randomly assigned treatments, compared with a placebo group, repeated across many patients, and possibly blocked by age.</p>

<p><strong>(iii)</strong> Three graphs used for quantitative variables are:
histogram, boxplot, and time series plot.</p>

<p><strong>(iv)</strong> An outlier is an observation that lies unusually far from the rest of the data values.</p>

<p><strong>(v)</strong> If the units are metres:
mean = metres,
median = metres,
standard deviation = metres,
variance = metres² (squared units).</p>

<p><strong>(vi)</strong>
For skewed distributions: use median and IQR.
For symmetric distributions: use mean and standard deviation.</p>

</div>

</div>

<script>
(function(){
  const show = document.getElementById("revq-show");
  const hide = document.getElementById("revq-hide");
  const model = document.getElementById("revq-model");
  const reset = document.getElementById("revq-reset");

  show.onclick = () => { model.style.display="block"; show.style.display="none"; hide.style.display="inline-block"; };
  hide.onclick = () => { model.style.display="none"; hide.style.display="none"; show.style.display="inline-block"; };
  
  reset.onclick = () => {
    document.querySelectorAll("#revq textarea").forEach(t => t.value = "");
    document.querySelectorAll("#revq input[type=checkbox]").forEach(c => c.checked = false);
    model.style.display="none";
    hide.style.display="none";
    show.style.display="inline-block";
  };
})();
</script>
```

<!-- PDF VERSION -->
```{r, results='asis', echo=FALSE}
if (knitr::is_latex_output()) {
  cat('
\\textit{Write your answers in the spaces provided.}

\\medskip

\\textbf{(i) What are the main differences between an observational study and an experiment?}
\\begin{answerbox}\\vspace{3cm}\\end{answerbox}

\\textbf{(ii) State and define, with examples, the four principles of experimental design.}
\\begin{answerbox}\\vspace{3cm}\\end{answerbox}

\\textbf{(iii) Which three types of graphs can be used to display the distribution of one quantitative variable?}
\\begin{answerbox}\\vspace{3cm}\\end{answerbox}

\\textbf{(iv) What is an outlier?}
\\begin{answerbox}\\vspace{3cm}\\end{answerbox}

\\textbf{(v) If the units are metres, what are the units of mean, median, standard deviation and variance?}
\\begin{answerbox}\\vspace{3cm}\\end{answerbox}

\\textbf{(vi) Which descriptions of centre and spread are best for skewed vs symmetric distributions?}
\\begin{answerbox}\\vspace{3cm}\\end{answerbox}
')
}
```

<!--
NOTE FOR FUTURE ME:
This module uses child Rmd files under section headings (##) so that:
- Module 1, Tutorial are in different files but render together
- Summary, Review, and Tutorial are numbered 1.1, 1.2, 1.3
- bs4_book splits these sections into separate HTML pages (via split_by: section)
- PDF output still renders cleanly with correct numbering

DO NOT add this file's children to _bookdown.yml
DO NOT use # headings in the child files
-->


```{r, child="02-Tut2.Rmd"}
```

