# Tutorial Questions: Module 2 {-}

Note: * Indicates **Expanded** Competencies

## Question 4: Describing Distributions

The following represent the assignment results (out of 100) for a class of 20 Accounting students:

85, 38, 92, 93, 100, 88, 78, 79, 96, 81, 62, 39, 77, 78, 83, 91, 90, 85, 99, 97

```{r hist-assignment, fig.cap="Histogram of assignment results (n = 20)", echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)

scores <- c(85, 38, 92, 93, 100, 88, 78, 79, 96, 81, 62, 39, 77, 78, 83, 91, 90, 85, 99, 97)

p <- ggplot(data.frame(scores), aes(x = scores)) +
  geom_histogram(binwidth = 10, fill = "#2a7ae2", color = "white", boundary = 0) +
  labs(title = "Histogram of Assignment Results (n=20)",
       x = "Assignment Results (out of 100)",
       y = "Frequency") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12)
  )

if (knitr::is_html_output()) {
  library(plotly)
  ggplotly(p)
} else {
  print(p)
}
```

For this data and Figure \@ref(fig:hist-assignment) complete the following questions:

<!-- HTML -->
```{=html}
<style>
  .q4-box { max-width: 750px; margin: 0; padding: 0; }
  .q4-btn { padding: 6px 12px; margin-right: 6px; margin-top: 10px; }
  .q4-hint {
    display: none;
    border-left: 6px solid #2b8cbe;
    background: #f6fbff;
    padding: .8rem 1rem;
    border-radius: 6px;
    margin-top: .6rem;
  }
  select.q4-correct, input.q4-correct, textarea.q4-correct { border: 3px solid green; background: #eaffea; }
  select.q4-incorrect, input.q4-incorrect, textarea.q4-incorrect { border: 3px solid red; background: #ffecec; }
</style>

<div class="q4-box">

<!-- ========================= Q4 (a) ========================= -->
<p><strong>(a)</strong> Describe the distribution of Assignment Results in terms of shape, centre, spread and outliers (including how many student results could be considered to be outliers).</p>

<p><strong>Shape:</strong></p>
<select id="q4-shape">
  <option value="">-- Select --</option>
  <option>Symmetric</option>
  <option>Skewed left, unimodal</option>
  <option>Skewed right, unimodal</option>
  <option>Skewed left, bimodal</option>
  <option>Skewed right, bimodal</option>
</select>
<div>
  <button class="q4-btn" onclick="q4ToggleHint('q4-hint-shape')">Show hint</button>
  <button class="q4-btn" onclick="q4CheckSelect('q4-shape','Skewed left, unimodal','q4-fb-shape')">Check</button>
  <button class="q4-btn" onclick="q4ResetSelect('q4-shape','q4-fb-shape','q4-hint-shape')">Reset</button>
</div>
<div class="q4-hint" id="q4-hint-shape">Look for a long tail towards the lower values. Note the two very low scores (below 40) far from the rest.</div>
<p id="q4-fb-shape"></p>

<p><strong>Centre:</strong></p>
<select id="q4-centre">
  <option value="">-- Select --</option>
  <option>60–69</option>
  <option>70–79</option>
  <option>80–89</option>
  <option>90–99</option>
</select>
<div>
  <button class="q4-btn" onclick="q4ToggleHint('q4-hint-centre')">Show hint</button>
  <button class="q4-btn" onclick="q4CheckSelect('q4-centre','80–89','q4-fb-centre')">Check</button>
  <button class="q4-btn" onclick="q4ResetSelect('q4-centre','q4-fb-centre','q4-hint-centre')">Reset</button>
</div>
<div class="q4-hint" id="q4-hint-centre">Most values cluster in the 80s — look at where the tallest bars of the histogram fall.</div>
<p id="q4-fb-centre"></p>

<p><strong>Spread:</strong></p>
<select id="q4-spread">
  <option value="">-- Select --</option>
  <option>38–100</option>
  <option>62–100</option>
  <option>70–95</option>
</select>
<div>
  <button class="q4-btn" onclick="q4ToggleHint('q4-hint-spread')">Show hint</button>
  <button class="q4-btn" onclick="q4CheckSelect('q4-spread','62–100','q4-fb-spread')">Check</button>
  <button class="q4-btn" onclick="q4ResetSelect('q4-spread','q4-fb-spread','q4-hint-spread')">Reset</button>
</div>
<div class="q4-hint" id="q4-hint-spread">Report the range of the main cluster of data, excluding obvious outliers (the two very low scores).</div>
<p id="q4-fb-spread"></p>

<p><strong>Outliers:</strong></p>
<select id="q4-outlier">
  <option value="">-- Select --</option>
  <option>None</option>
  <option>1 student</option>
  <option>2 students</option>
</select>
<div>
  <button class="q4-btn" onclick="q4ToggleHint('q4-hint-outlier')">Show hint</button>
  <button class="q4-btn" onclick="q4CheckSelect('q4-outlier','2 students','q4-fb-outlier')">Check</button>
  <button class="q4-btn" onclick="q4ResetSelect('q4-outlier','q4-fb-outlier','q4-hint-outlier')">Reset</button>
</div>
<div class="q4-hint" id="q4-hint-outlier">How many values are clearly separated from the main cluster of data? Look for values below 60.</div>
<p id="q4-fb-outlier"></p>

<hr>

<!-- ========================= Q4 (b) ========================= -->
<p><strong>(b)</strong> What centre and spread would you use to describe this distribution? Why?</p>
<input id="q4-b" style="width:100%;" placeholder="Type your answer here…">
<div>
  <button class="q4-btn" onclick="q4ToggleHint('q4-hint-b')">Show hint</button>
  <button class="q4-btn" onclick="q4CheckKeywords('q4-b',['median'],['iqr','interquartile'],'q4-fb-b','Correct ✅ — median and IQR are the appropriate measures for a skewed distribution with outliers.')">Check</button>
  <button class="q4-btn" onclick="q4ResetText('q4-b','q4-fb-b','q4-hint-b')">Reset</button>
</div>
<div class="q4-hint" id="q4-hint-b">If the distribution is skewed or has outliers, use the <strong>median</strong> (centre) and <strong>IQR</strong> (spread). If symmetric with no outliers, use the <strong>mean</strong> and <strong>standard deviation</strong>.</div>
<p id="q4-fb-b"></p>

<hr>

<!-- ========================= Q4 (c) ========================= -->
<p><strong>(c)</strong> Calculate the mean of the data set:</p>
<input type="number" id="q4-mean" placeholder="Enter mean…">
<div>
  <button class="q4-btn" onclick="q4ToggleHint('q4-hint-c')">Show hint</button>
  <button class="q4-btn" onclick="q4CheckNum('q4-mean',81.55,'q4-fb-c')">Check</button>
  <button class="q4-btn" onclick="q4ResetText('q4-mean','q4-fb-c','q4-hint-c')">Reset</button>
</div>
<div class="q4-hint" id="q4-hint-c">Add all 20 values together, then divide by 20. The sum is 1631.</div>
<p id="q4-fb-c"></p>

<hr>

<!-- ========================= Q4 (d) ========================= -->
<p><strong>(d)</strong> Calculate the standard deviation of the data set:</p>
<input type="number" id="q4-sd" placeholder="Enter SD…">
<div>
  <button class="q4-btn" onclick="q4ToggleHint('q4-hint-d')">Show hint</button>
  <button class="q4-btn" onclick="q4CheckNum('q4-sd',17.32,'q4-fb-d')">Check</button>
  <button class="q4-btn" onclick="q4ResetText('q4-sd','q4-fb-d','q4-hint-d')">Reset</button>
</div>
<div class="q4-hint" id="q4-hint-d">Use the <strong>sample</strong> standard deviation formula (divide by n − 1 = 19). On most calculators this is the <em>s</em> or <em>s<sub>x</sub></em> button.</div>
<p id="q4-fb-d"></p>

<hr>

<!-- ========================= Q4 (e) ========================= -->
<p><strong>(e)</strong> Calculate the median of the data set by hand:</p>
<input type="number" id="q4-median" placeholder="Enter median…">
<div>
  <button class="q4-btn" onclick="q4ToggleHint('q4-hint-e')">Show hint</button>
  <button class="q4-btn" onclick="q4CheckNum('q4-median',85,'q4-fb-e')">Check</button>
  <button class="q4-btn" onclick="q4ResetText('q4-median','q4-fb-e','q4-hint-e')">Reset</button>
</div>
<div class="q4-hint" id="q4-hint-e">Sort all 20 values from smallest to largest. With an even number of values, the median is the average of the 10th and 11th values in the sorted list.</div>
<p id="q4-fb-e"></p>

<hr>

<!-- ========================= Q4 (f) ========================= -->
<p><strong>(f)</strong> Why are the mean and median different/similar?</p>
<textarea id="q4-f" style="width:100%;" rows="4" placeholder="Type your explanation here…"></textarea>
<div>
  <button class="q4-btn" onclick="q4ToggleHint('q4-hint-f')">Show hint</button>
  <button class="q4-btn" id="q4-model-f-show" onclick="q4ToggleHint('q4-model-f')">Show model answer</button>
  <button class="q4-btn" onclick="q4CheckKeywords('q4-f',['mean','median'],['outlier','skew','pull','different'],'q4-fb-f','Good explanation ✅')">Check</button>
  <button class="q4-btn" onclick="q4ResetText('q4-f','q4-fb-f','q4-hint-f')">Reset</button>
</div>
<div class="q4-hint" id="q4-hint-f">Think about the effect of extreme values (outliers). Which measure of centre is pulled toward very low or very high scores?</div>
<div class="q4-hint" id="q4-model-f">The mean (81.55) and median (85) are different because the two very low scores (38 and 39) pull the mean downward. Outliers have a strong influence on the mean but very little effect on the median. As a result, in this left-skewed distribution the median is noticeably higher than the mean.</div>
<p id="q4-fb-f"></p>

<hr>

<!-- ========================= Q4 (g) ========================= -->
<p><strong>(g)</strong> Reproduce the histogram and descriptive statistics using statistical software (e.g. Excel or SPSS) and compare your results with your hand calculations above.</p>

<button class="q4-btn" onclick="q4ToggleHint('q4-hint-g')">Show hint</button>
<div class="q4-hint" id="q4-hint-g">
  <strong>Excel:</strong> Enter the data in a column. Use <em>Data → Data Analysis → Histogram</em> for the histogram and <em>Data → Data Analysis → Descriptive Statistics</em> for summary statistics (tick "Summary statistics").<br><br>
  <strong>SPSS:</strong> Use <em>Analyze → Descriptive Statistics → Frequencies</em>, click <em>Statistics</em> to tick mean, median, SD etc., and click <em>Charts</em> to request a histogram.
</div>

<hr>

<!-- ========================= Q4 (h) ========================= -->
<p><strong>(h)</strong> What five numbers make up the five-number summary?</p>
<input id="q4-h" style="width:100%;" placeholder="Name the five numbers (e.g. minimum, Q1, …)">
<div>
  <button class="q4-btn" onclick="q4ToggleHint('q4-hint-h')">Show hint</button>
  <button class="q4-btn" onclick="q4CheckFiveNum()">Check</button>
  <button class="q4-btn" onclick="q4ResetText('q4-h','q4-fb-h','q4-hint-h')">Reset</button>
</div>
<div class="q4-hint" id="q4-hint-h">The five-number summary uses five key statistics that together summarise the full distribution. Think: one at each end, one in the middle, and one in each half.</div>
<p id="q4-fb-h"></p>

<hr>

<!-- ========================= Q4 (i) ========================= -->
<p><strong>(i)</strong> Calculate Q1 and Q3:</p>
<p>Q1: <input type="number" id="q4-q1" placeholder="Q1…" style="width:120px;"></p>
<p>Q3: <input type="number" id="q4-q3" placeholder="Q3…" style="width:120px;"></p>
<div>
  <button class="q4-btn" onclick="q4ToggleHint('q4-hint-i')">Show hint</button>
  <button class="q4-btn" onclick="q4CheckI()">Check</button>
  <button class="q4-btn" onclick="q4ResetI()">Reset</button>
</div>
<div class="q4-hint" id="q4-hint-i">
  Sort the 20 values from smallest to largest. Split them into a lower half (positions 1–10) and an upper half (positions 11–20).<br>
  Q1 = median of the lower half = average of the 5th and 6th values.<br>
  Q3 = median of the upper half = average of the 15th and 16th values.
</div>
<p id="q4-fb-i"></p>

<hr>

<!-- ========================= Q4 (j) ========================= -->
<p><strong>(j)</strong> Calculate the IQR:</p>
<input type="number" id="q4-iqr" placeholder="IQR…">
<div>
  <button class="q4-btn" onclick="q4ToggleHint('q4-hint-j')">Show hint</button>
  <button class="q4-btn" onclick="q4CheckNum('q4-iqr',14.5,'q4-fb-j')">Check</button>
  <button class="q4-btn" onclick="q4ResetText('q4-iqr','q4-fb-j','q4-hint-j')">Reset</button>
</div>
<div class="q4-hint" id="q4-hint-j">IQR = Q3 − Q1. Use your answers from part (i).</div>
<p id="q4-fb-j"></p>

</div><!-- end .q4-box -->

<!-- ========================= JAVASCRIPT ========================= -->
<script>
(function(){

  var TOL = 0.11;

  /* ------ Hint toggle ------ */
  window.q4ToggleHint = function(id){
    var el = document.getElementById(id);
    if(!el) return;
    el.style.display = (el.style.display === "block") ? "none" : "block";
  };

  /* ------ Dropdown check ------ */
  window.q4CheckSelect = function(selId, correct, fbId){
    var el = document.getElementById(selId);
    var fb = document.getElementById(fbId);
    var ok = (el.value === correct);
    el.className = ok ? "q4-correct" : "q4-incorrect";
    fb.innerHTML = ok
      ? "<span style='color:green;'>Correct \u2705</span>"
      : "<span style='color:red;'>Not quite \u2014 use the hint if needed.</span>";
    localStorage.setItem(selId, el.value);
  };

  window.q4ResetSelect = function(selId, fbId, hintId){
    var el = document.getElementById(selId);
    el.value = ""; el.className = "";
    document.getElementById(fbId).innerHTML = "";
    var h = document.getElementById(hintId);
    if(h) h.style.display = "none";
    localStorage.removeItem(selId);
  };

  /* ------ Numeric check ------ */
  window.q4CheckNum = function(id, expected, fbId){
    var el = document.getElementById(id);
    var fb = fbId ? document.getElementById(fbId) : null;
    var v  = parseFloat(el.value);
    var ok = !isNaN(v) && Math.abs(v - expected) <= TOL;
    el.className = ok ? "q4-correct" : "q4-incorrect";
    if(fb) fb.innerHTML = ok
      ? "<span style='color:green;'>Correct \u2705</span>"
      : "<span style='color:red;'>Not quite \u2014 check your working and use the hint if needed.</span>";
    localStorage.setItem(id, el.value);
  };

  /* ------ Keyword check (text / textarea) ------ */
  window.q4CheckKeywords = function(id, required, any, fbId, successMsg){
    var el = document.getElementById(id);
    var fb = document.getElementById(fbId);
    var t  = el.value.toLowerCase();
    var hasRequired = required.every(function(w){ return t.includes(w); });
    var hasAny      = any.some(function(w){ return t.includes(w); });
    var ok = hasRequired && hasAny;
    el.className = ok ? "q4-correct" : "q4-incorrect";
    fb.innerHTML = ok
      ? "<span style='color:green;'>" + successMsg + "</span>"
      : "<span style='color:red;'>Make sure your answer mentions the key terms \u2014 try the hint.</span>";
    localStorage.setItem(id, el.value);
  };

  /* ------ Generic text/number reset ------ */
  window.q4ResetText = function(id, fbId, hintId){
    var el = document.getElementById(id);
    el.value = ""; el.className = "";
    if(fbId){ document.getElementById(fbId).innerHTML = ""; }
    var h = document.getElementById(hintId);
    if(h) h.style.display = "none";
    localStorage.removeItem(id);
  };

  /* ------ (f) model answer independent toggle (already handled by q4ToggleHint) ------ */

  /* ------ (h) Five-number summary ------ */
  window.q4CheckFiveNum = function(){
    var el = document.getElementById("q4-h");
    var fb = document.getElementById("q4-fb-h");
    var t  = el.value.toLowerCase();
    var ok = t.includes("min") && t.includes("q1") && t.includes("median") &&
             t.includes("q3") && t.includes("max");
    el.className = ok ? "q4-correct" : "q4-incorrect";
    fb.innerHTML = ok
      ? "<span style='color:green;'>Correct \u2705</span>"
      : "<span style='color:red;'>Make sure you name all five: minimum, Q1, median, Q3, maximum.</span>";
    localStorage.setItem("q4-h", el.value);
  };

  /* ------ (i) Q1 and Q3 ------ */
  window.q4CheckI = function(){
    var elQ1 = document.getElementById("q4-q1");
    var elQ3 = document.getElementById("q4-q3");
    var fb   = document.getElementById("q4-fb-i");
    var okQ1 = !isNaN(parseFloat(elQ1.value)) && Math.abs(parseFloat(elQ1.value) - 78)   <= TOL;
    var okQ3 = !isNaN(parseFloat(elQ3.value)) && Math.abs(parseFloat(elQ3.value) - 92.5) <= TOL;
    elQ1.className = okQ1 ? "q4-correct" : "q4-incorrect";
    elQ3.className = okQ3 ? "q4-correct" : "q4-incorrect";
    fb.innerHTML = (okQ1 && okQ3)
      ? "<span style='color:green;'>Both correct \u2705 (Q1 = 78, Q3 = 92.5)</span>"
      : "<span style='color:red;'>Check: Q1 = 78 and Q3 = 92.5. Use the hint for the method.</span>";
    localStorage.setItem("q4-q1", elQ1.value);
    localStorage.setItem("q4-q3", elQ3.value);
  };

  window.q4ResetI = function(){
    ["q4-q1","q4-q3"].forEach(function(id){
      var el = document.getElementById(id);
      if(el){ el.value = ""; el.className = ""; }
      localStorage.removeItem(id);
    });
    document.getElementById("q4-fb-i").innerHTML = "";
    var h = document.getElementById("q4-hint-i");
    if(h) h.style.display = "none";
  };

  /* ------ Restore saved answers on page load ------ */
  var allIds = [
    "q4-shape","q4-centre","q4-spread","q4-outlier",
    "q4-b","q4-mean","q4-sd","q4-median","q4-f",
    "q4-h","q4-q1","q4-q3","q4-iqr"
  ];
  allIds.forEach(function(id){
    var saved = localStorage.getItem(id);
    if(saved){
      var el = document.getElementById(id);
      if(el) el.value = saved;
    }
  });

})();
</script>
```

<!-- PDF -->
```{r, results='asis', echo=FALSE}
if (knitr::is_latex_output()) {
cat('
\\textbf{(a) Describe the distribution (shape, centre, spread and outliers).}
\\begin{answerbox}\\vspace{4cm}\\end{answerbox}

\\textbf{(b) What centre and spread would you use? Why?}
\\begin{answerbox}\\vspace{3cm}\\end{answerbox}

\\textbf{(c) Calculate the mean.}
\\begin{answerbox}\\vspace{2cm}\\end{answerbox}

\\textbf{(d) Calculate the standard deviation.}
\\begin{answerbox}\\vspace{2cm}\\end{answerbox}

\\textbf{(e) Calculate the median.}
\\begin{answerbox}\\vspace{2cm}\\end{answerbox}

\\textbf{(f) Why are the mean and median different/similar?}
\\begin{answerbox}\\vspace{3cm}\\end{answerbox}

\\textbf{(g) Reproduce the histogram and descriptive statistics in software.}
\\begin{answerbox}\\vspace{3cm}\\end{answerbox}

\\textbf{(h) What are the five numbers in the five-number summary?}
\\begin{answerbox}\\vspace{2cm}\\end{answerbox}

\\textbf{(i) Calculate Q1 and Q3.}
\\begin{answerbox}\\vspace{2cm}\\end{answerbox}

\\textbf{(j) Calculate the IQR.}
\\begin{answerbox}\\vspace{2cm}\\end{answerbox}
')
}
```


## Further Practice from the Textbook

Answers to these questions are located in the back of the textbook. If you require further explanation, please ask in the Module Content Discussion Forum on the StudyDesk.

- Questions 3.25, 4.31, 4.33 (De Veaux, Velleman & Bock, 4th edition)

- Questions 2.53, 4.31, 4.33 (De Veaux, Velleman & Bock, 5th edition)

## Textbook Readings

For extra clarification on materials in this module, see: De Veaux, Velleman & Bock (4th edition), Chapters 3, 4, and 12 (De Veaux, Velleman & Bock (5th edition), Chapters 2.2--2.5, 4, and 11).


# Module 2 Checklist {-}

```{=html}
<style>
  .cl-wrap { max-width: 900px; margin: 0 auto; }
  .cl-card { border: 1px solid #e5e5e5; border-radius: 6px; padding: 14px; background: #fff; margin: 12px 0; }
  .cl-card h4 { margin: .2rem 0 .6rem; font-size: 1.05rem; }
  .cl-note { font-size: .9rem; color: #666; }
  .cl-group-title { font-weight: 700; margin: .4rem 0; }
  .cl-item { display: flex; align-items: flex-start; gap: 8px; margin: .35rem 0; }
  .cl-item input[type=checkbox] { transform: scale(1.2); margin-top: 2px; }
  .cl-adv { font-style: italic; color: #333; }
  .cl-controls { display: flex; gap: 8px; flex-wrap: wrap; margin-top: .6rem; }
  .btn { padding: 6px 10px; border: 1px solid #ccc; border-radius: 4px; background: #f7f7f7; cursor: pointer; }
  .cl-prog { height: 10px; background: #f1f1f1; border-radius: 6px; overflow: hidden; margin: .4rem 0 .8rem; }
  .cl-bar { height: 100%; width: 0%; background: #2b8cbe; transition: width .25s ease; }
  .cl-stats { font-size: .9rem; color: #555; }
</style>

<div class="cl-wrap" id="m2-checklist">
  <div class="cl-card">
    <h4>Module 2 — Checklist</h4>
    <div class="cl-note">
      Once you have watched the lecture recordings, completed the tutorial activity,
      and textbook problems, tick the outcomes you have achieved.
      Your selections are saved in your browser on this device.
    </div>
    <div class="cl-prog"><div class="cl-bar"></div></div>
    <div class="cl-stats"></div>
    <div class="cl-controls">
      <button class="btn" data-action="all">Mark all done</button>
      <button class="btn" data-action="clear">Clear all</button>
      <span class="cl-note">(* denotes expanded competencies)</span>
    </div>
  </div>

  <div class="cl-card">
    <div class="cl-group-title">Describing Distributions</div>

    <label class="cl-item">
      <input type="checkbox" data-id="d1">
      <span>Visually describe the distribution of a quantitative variable in terms of shape, centre, and spread from a histogram.</span>
    </label>
    <label class="cl-item">
      <input type="checkbox" data-id="d2">
      <span>Calculate by hand and interpret the median.</span>
    </label>
    <label class="cl-item">
      <input type="checkbox" data-id="d3">
      <span>Know the numbers that make up the five-number summary.</span>
    </label>
    <label class="cl-item">
      <input type="checkbox" data-id="d4">
      <span>Use technology to find the mean and standard deviation of a data set.</span>
    </label>
    <label class="cl-item">
      <input type="checkbox" data-id="d5">
      <span>Compare groups using boxplots.</span>
    </label>
    <label class="cl-item">
      <input type="checkbox" data-id="d6">
      <span class="cl-adv">* Calculate by hand the interquartile range (IQR) and five-number summary.</span>
    </label>
    <label class="cl-item">
      <input type="checkbox" data-id="d7">
      <span class="cl-adv">* Recognise the impact of outliers and skewness on measures of centre and spread.</span>
    </label>
  </div>
</div>

<script>
(function(){
  var ROOT = document.getElementById("m2-checklist");
  var KEY  = "m2-checklist-v1";
  var boxes = Array.from(ROOT.querySelectorAll("input[type=checkbox][data-id]"));
  var bar   = ROOT.querySelector(".cl-bar");
  var stats = ROOT.querySelector(".cl-stats");
  var btnAll   = ROOT.querySelector('[data-action="all"]');
  var btnClear = ROOT.querySelector('[data-action="clear"]');

  function update(){
    var total = boxes.length;
    var done  = boxes.filter(function(b){ return b.checked; }).length;
    var pct   = total ? Math.round(100 * done / total) : 0;
    bar.style.width = pct + "%";
    stats.textContent = done + " of " + total + " completed (" + pct + "%)";
  }
  function save(){
    var state = {};
    boxes.forEach(function(b){ state[b.dataset.id] = b.checked; });
    localStorage.setItem(KEY, JSON.stringify(state));
    update();
  }
  function load(){
    var saved = JSON.parse(localStorage.getItem(KEY) || "{}");
    boxes.forEach(function(b){ b.checked = !!saved[b.dataset.id]; });
    update();
  }
  boxes.forEach(function(b){ b.addEventListener("change", save); });
  btnAll.addEventListener("click", function(){ boxes.forEach(function(b){ b.checked = true; }); save(); });
  btnClear.addEventListener("click", function(){
    if(!confirm("Clear all ticks for this checklist on this device?")) return;
    localStorage.removeItem(KEY);
    boxes.forEach(function(b){ b.checked = false; });
    update();
  });
  load();
})();
</script>
```

## Using Statistical Software Checklist {-}

```{=html}
<div class="cl-wrap" id="m2-software-checklist">
  <div class="cl-card">
    <h4>Module 2 — Using Statistical Software Checklist</h4>
    <div class="cl-note">
      On successful completion of this module, students should be able to use SPSS or Excel
      to complete the following tasks.
    </div>
    <div class="cl-prog"><div class="cl-bar"></div></div>
    <div class="cl-stats"></div>
    <div class="cl-controls">
      <button class="btn" data-action="all">Mark all done</button>
      <button class="btn" data-action="clear">Clear all</button>
    </div>
  </div>

  <div class="cl-card">
    <div class="cl-group-title">Using Statistical Software</div>
    <label class="cl-item">
      <input type="checkbox" data-id="ss1">
      <span>Construct a histogram, a boxplot, and multiple boxplots.</span>
    </label>
    <label class="cl-item">
      <input type="checkbox" data-id="ss2">
      <span>Produce summary statistics of quantitative data.</span>
    </label>
    <label class="cl-item">
      <input type="checkbox" data-id="ss3">
      <span>Select subsets of data.</span>
    </label>
  </div>
</div>

<script>
(function(){
  var ROOT = document.getElementById("m2-software-checklist");
  var KEY  = "m2-software-checklist-v1";
  var boxes = Array.from(ROOT.querySelectorAll("input[type=checkbox][data-id]"));
  var bar   = ROOT.querySelector(".cl-bar");
  var stats = ROOT.querySelector(".cl-stats");
  var btnAll   = ROOT.querySelector('[data-action="all"]');
  var btnClear = ROOT.querySelector('[data-action="clear"]');

  function update(){
    var total = boxes.length;
    var done  = boxes.filter(function(b){ return b.checked; }).length;
    var pct   = total ? Math.round(100 * done / total) : 0;
    bar.style.width = pct + "%";
    stats.textContent = done + " of " + total + " completed (" + pct + "%)";
  }
  function save(){
    var state = {};
    boxes.forEach(function(b){ state[b.dataset.id] = b.checked; });
    localStorage.setItem(KEY, JSON.stringify(state));
    update();
  }
  function load(){
    var saved = JSON.parse(localStorage.getItem(KEY) || "{}");
    boxes.forEach(function(b){ b.checked = !!saved[b.dataset.id]; });
    update();
  }
  boxes.forEach(function(b){ b.addEventListener("change", save); });
  btnAll.addEventListener("click", function(){ boxes.forEach(function(b){ b.checked = true; }); save(); });
  btnClear.addEventListener("click", function(){
    if(!confirm("Clear all ticks for this checklist on this device?")) return;
    localStorage.removeItem(KEY);
    boxes.forEach(function(b){ b.checked = false; });
    update();
  });
  load();
})();
</script>
```

<!-- PDF checklist -->
```{r, echo=FALSE, results='asis'}
if (knitr::is_latex_output()) {
  cat("
\\section*{Describing Distributions}

- \\(\\square\\) Visually describe the distribution of a quantitative variable in terms of shape, centre, and spread from a histogram
- \\(\\square\\) Calculate by hand and interpret the median
- \\(\\square\\) Know the numbers that make up the five-number summary
- \\(\\square\\) Use technology to find the mean and standard deviation of a data set
- \\(\\square\\) Compare groups using boxplots
- \\(\\square\\) \\emph{*Calculate by hand the interquartile range (IQR) and five-number summary}
- \\(\\square\\) \\emph{*Recognise the impact of outliers and skewness on measures of centre and spread}

\\section*{Using Statistical Software}

- \\(\\square\\) Construct a histogram, a boxplot, and multiple boxplots
- \\(\\square\\) Produce summary statistics of quantitative data
- \\(\\square\\) Select subsets of data using Select Cases
")
}
```
