# Tutorial Questions: Module 2 {-}

Note: * Indicates **Expanded** Competencies

## Question 4: Describing Distributions

The following represent the assignment results (out of 100) for a class of 20 Accounting students:

85, 38, 92, 93, 100, 88, 78, 79, 96, 81, 62, 39, 77, 78, 83, 91, 90, 85, 99, 97

```{r hist-assignment, fig.cap="Histogram of assignment results (n = 20)", echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(plotly)

# Data
scores <- c(85, 38, 92, 93, 100, 88, 78, 79, 96, 81, 62, 39, 77, 78, 83, 91, 90, 85, 99, 97)

# Histogram
p <- ggplot(data.frame(scores), aes(x = scores)) +
  geom_histogram(binwidth = 10, fill = "#2a7ae2", color = "white", boundary = 0) +
  labs(title = "Histogram of Assignment Results (n=20)",
       x = "Assignment Results (out of 100)",
       y = "Frequency") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12)
  )

# Interactive
ggplotly(p)
```


For this data and Figure \@ref(fig:hist-assignment) complete the following questions:
<!-- HTML -->
```{=html}

<!-- =========================
Q4 (a) Describe the distribution
========================== -->
<p><strong>(a)</strong> Describe the distribution of Assignment Results in terms of shape, centre, spread and outliers (including how many student results could be considered to be outliers).</p>
<p><strong>Shape:</strong></p>
<select id="q4-shape">
  <option value="">Choose</option>
  <option>Symmetric</option>
  <option>Skewed left, unimodal</option>
  <option>Skewed right, unimodal</option>
  <option>Skewed left, bimodal</option>
  <option>Skewed right, bimodal</option>
</select>
<div>
  <button class="q4-btn" onclick="q4_hintA('shape')">Hint</button>
  <button class="q4-btn" onclick="q4_checkA('shape')">Check</button>
  <button class="q4-btn" onclick="q4_resetA('shape')">Reset</button>
</div>
<p id="q4-fb-shape"></p>

<p><strong>Centre:</strong></p>
<select id="q4-centre">
  <option value="">Choose</option>
  <option>60–69</option>
  <option>70–79</option>
  <option>80–89</option>
  <option>90–99</option>
</select>
<div>
  <button class="q4-btn" onclick="q4_hintA('centre')">Hint</button>
  <button class="q4-btn" onclick="q4_checkA('centre')">Check</button>
  <button class="q4-btn" onclick="q4_resetA('centre')">Reset</button>
</div>
<p id="q4-fb-centre"></p>

<p><strong>Spread:</strong></p>
<select id="q4-spread">
  <option value="">Choose</option>
  <option>38–100</option>
  <option>62–100</option>
  <option>70–95</option>
</select>
<div>
  <button class="q4-btn" onclick="q4_hintA('spread')">Hint</button>
  <button class="q4-btn" onclick="q4_checkA('spread')">Check</button>
  <button class="q4-btn" onclick="q4_resetA('spread')">Reset</button>
</div>
<p id="q4-fb-spread"></p>

<p><strong>Outliers:</strong></p>
<select id="q4-outlier">
  <option value="">Choose</option>
  <option>None</option>
  <option>1 student</option>
  <option>2 students</option>
</select>
<div>
  <button class="q4-btn" onclick="q4_hintA('outlier')">Hint</button>
  <button class="q4-btn" onclick="q4_checkA('outlier')">Check</button>
  <button class="q4-btn" onclick="q4_resetA('outlier')">Reset</button>
</div>
<p id="q4-fb-outlier"></p>

<hr>

<!-- =========================
(b) Centre and spread
========================== -->
<p><strong>(b)</strong> What centre and spread would you use to describe this distribution? Why?</p>
<input id="q4-b" style="width:100%;">
<div>
  <button onclick="q4_hintB()">Hint</button>
  <button onclick="q4_checkB()">Check</button>
  <button onclick="q4_resetB()">Reset</button>
</div>
<p id="q4-fb-b"></p>

<hr>

<!-- =========================
(c) Mean
========================== -->
<p><strong>(c)</strong> Calculate the mean of the data set:</p>
<input type="number" id="q4-mean">
<div>
  <button onclick="q4_hintC()">Hint</button>
  <button onclick="q4_checkC()">Check</button>
  <button onclick="q4_resetC()">Reset</button>
</div>
<p id="q4-fb-c"></p>

<hr>

<!-- =========================
(d) SD
========================== -->
<p><strong>(d)</strong> Calculate the standard deviation of the data set:</p>
<input type="number" id="q4-sd">
<div>
  <button onclick="q4_hintD()">Hint</button>
  <button onclick="q4_checkD()">Check</button>
  <button onclick="q4_resetD()">Reset</button>
</div>
<p id="q4-fb-d"></p>

<hr>

<!-- =========================
(e) Median
========================== -->
<p><strong>(e)</strong> Calculate the median of the data set by hand:</p>
<input type="number" id="q4-median">
<div>
  <button onclick="q4_hintE()">Hint</button>
  <button onclick="q4_checkE()">Check</button>
  <button onclick="q4_resetE()">Reset</button>
</div>
<p id="q4-fb-e"></p>

<hr>

<!-- =========================
(f) Explanation
========================== -->
<p><strong>(f)</strong> Why are the mean and median different/similar?</p>
<textarea id="q4-f" style="width:100%;"></textarea>
<div>
  <button onclick="q4_hintF()">Hint</button>
  <button onclick="q4_checkF()">Check</button>
  <button onclick="q4_resetF()">Reset</button>
</div>
<p id="q4-fb-f"></p>

<hr>

<!-- =========================
(h) Five-number summary
========================== -->
<p><strong>(h)</strong> What five numbers make up the five-number summary?</p>
<input id="q4-h" style="width:100%;">
<div>
  <button onclick="q4_hintH()">Hint</button>
  <button onclick="q4_checkH()">Check</button>
  <button onclick="q4_resetH()">Reset</button>
</div>
<p id="q4-fb-h"></p>

<hr>

<!-- =========================
(i) Quartiles
========================== -->
<p><strong>(i)</strong> Q1:</p>
<input type="number" id="q4-q1">

<p>Q3:</p>
<input type="number" id="q4-q3">

<div>
  <button onclick="q4_hintI()">Hint</button>
  <button onclick="q4_checkI()">Check</button>
  <button onclick="q4_resetI()">Reset</button>
</div>
<p id="q4-fb-i"></p>

<hr>

<!-- =========================
(j) IQR
========================== -->
<p><strong>(j)</strong> IQR:</p>
<input type="number" id="q4-iqr">
<div>
  <button onclick="q4_hintJ()">Hint</button>
  <button onclick="q4_checkJ()">Check</button>
  <button onclick="q4_resetJ()">Reset</button>
</div>
<p id="q4-fb-j"></p>

<!-- =========================
LOGIC
========================== -->
<script>

const tol = 0.1;

function setState(id, ok){
  const el = document.getElementById(id);
  el.style.background = ok ? "#c8f7c5" : "#f7c5c5";
}

/* (a) */
function q4_checkA(x){
  const ans = {shape:"Skewed left, unimodal",centre:"80–89",spread:"62–100",outlier:"2 students"};
  const el = document.getElementById("q4-"+x);
  const ok = el.value === ans[x];
  setState("q4-"+x,ok);
}
/* hints */
function q4_hintA(x){
  const msg={
    shape:"Look for long tail and note the low values less than 40",
    centre:"Most values are in the 80s",
    spread:"Ignore outliers and focus on the bulk of the data",
    outlier:"How many low values?"
  };
  document.getElementById("q4-fb-"+x).innerHTML = msg[x];
}

function q4_resetA(x){
  document.getElementById("q4-"+x).value="";
  document.getElementById("q4-"+x).style.background="";
}

/* (b) */
function q4_checkB(){
  let t = document.getElementById("q4-b").value.toLowerCase();
  setState("q4-b", t.includes("median") && t.includes("iqr"));
}
function q4_hintB(){
  document.getElementById("q4-fb-b").innerHTML="Skewed or outliers then use median & IQR. If Symmetric then use mean & Standard deviation";
}
function q4_resetB(){
  document.getElementById("q4-b").value="";
  document.getElementById("q4-b").style.background="";
}

/* (c,d,e) */
function checkNum(id,val){
  let x = parseFloat(document.getElementById(id).value);
  setState(id, Math.abs(x-val)<=tol);
}
function q4_checkC(){ checkNum("q4-mean",81.55); }
function q4_checkD(){ checkNum("q4-sd",17.32); }
function q4_checkE(){ checkNum("q4-median",85); }

function q4_hintC(){document.getElementById("q4-fb-c").innerHTML="sum ÷ 20";}
function q4_hintD(){document.getElementById("q4-fb-d").innerHTML="use sample SD";}
function q4_hintE(){document.getElementById("q4-fb-e").innerHTML="average 10th & 11th";}

function resetNum(id){document.getElementById(id).value="";document.getElementById(id).style.background="";}
function q4_resetC(){resetNum("q4-mean");}
function q4_resetD(){resetNum("q4-sd");}
function q4_resetE(){resetNum("q4-median");}

/* (f) */
function q4_checkF(){
  let t=document.getElementById("q4-f").value.toLowerCase();
  setState("q4-f", t.includes("mean") && t.includes("median") && t.includes("outlier") && t.includes("different"));
}
function q4_hintF(){
  document.getElementById("q4-fb-f").innerHTML="Outliers pull mean.";
}
function q4_resetF(){
  document.getElementById("q4-f").value="";
}

/* (h) */
function q4_checkH(){
  let t=document.getElementById("q4-h").value.toLowerCase();
  setState("q4-h", t.includes("minimum") && t.includes("q1") && t.includes("median") && t.includes("q3") && t.includes("maximum"));
}
function q4_hintH(){
  document.getElementById("q4-fb-h").innerHTML="min, Q1, median, Q3, max";
}
function q4_resetH(){
  document.getElementById("q4-h").value="";
}

/* (i) */
function q4_checkI(){
  checkNum("q4-q1",78);
  checkNum("q4-q3",92.5);
}
function q4_hintI(){
  document.getElementById("q4-fb-i").innerHTML="Split ordered data into halves.";
}
function q4_resetI(){
  resetNum("q4-q1"); resetNum("q4-q3");
}

/* (j) */
function q4_checkJ(){ checkNum("q4-iqr",14.5); }
function q4_hintJ(){
  document.getElementById("q4-fb-j").innerHTML="IQR = Q3 − Q1";
}
function q4_resetJ(){ resetNum("q4-iqr"); }

</script>
```

<!-- PDF -->
```{r, results='asis', echo=FALSE}
if (knitr::is_latex_output()) {
cat('
\\textbf{(a) Describe the distribution (shape, centre, spread and outliers).}
\\begin{answerbox}\\vspace{4cm}\\end{answerbox}

\\textbf{(b) What centre and spread would you use? Why?}
\\begin{answerbox}\\vspace{3cm}\\end{answerbox}

\\textbf{(c) Calculate the mean.}
\\begin{answerbox}\\vspace{2cm}\\end{answerbox}

\\textbf{(d) Calculate the standard deviation.}
\\begin{answerbox}\\vspace{2cm}\\end{answerbox}

\\textbf{(e) Calculate the median.}
\\begin{answerbox}\\vspace{2cm}\\end{answerbox}

\\textbf{(f) Why are the mean and median different/similar?}
\\begin{answerbox}\\vspace{3cm}\\end{answerbox}

\\textbf{(g) Reproduce the histogram and descriptive statistics in software.}
\\begin{answerbox}\\vspace{3cm}\\end{answerbox}

\\textbf{(h) What are the five numbers in the five-number summary?}
\\begin{answerbox}\\vspace{2cm}\\end{answerbox}

\\textbf{(i) Calculate Q1 and Q3.}
\\begin{answerbox}\\vspace{2cm}\\end{answerbox}

\\textbf{(j) Calculate the IQR.}
\\begin{answerbox}\\vspace{2cm}\\end{answerbox}
')
}
```


## Further Practice from the Textbook

Answers to these questions are located in the back of the textbook. If you require further explanation, please ask in the Module Content Discussion Forum on the Study-
Desk.

• Questions 3.25, 4.31, 4.33  (De Veaux, Velleman & Bock, (4th edition))

• Questions 2.53, 4.31, 4.33  (De Veaux, Velleman & Bock, (5th edition))

## Textbook Readings 
For extra clarification on materials in this module, see: De Veaux, Velleman
& Bock, (4th edition), Chapters 3, 4, and 12 (De Veaux, Velleman
& Bock, (5th edition), Chapters 2.2-2.5, 4, and 11).


# Module 2 Checklist {-}

```{=html}
<style>
  .cl-wrap{max-width:900px;margin:0 auto}
  .cl-card{border:1px solid #e5e5e5;border-radius:6px;padding:14px;background:#fff;margin:12px 0}
  .cl-card h4{margin:.2rem 0 .6rem;font-size:1.05rem}
  .cl-note{font-size:.9rem;color:#666}
  .cl-group-title{font-weight:700;margin:.4rem 0}
  .cl-item{display:flex;align-items:flex-start;gap:8px;margin:.35rem 0}
  .cl-item input[type=checkbox]{transform:scale(1.2);margin-top:2px}
  .cl-adv{font-style:italic;color:#333}
  .cl-controls{display:flex;gap:8px;flex-wrap:wrap;margin-top:.6rem}
  .btn{padding:6px 10px;border:1px solid #ccc;border-radius:4px;background:#f7f7f7;cursor:pointer}
  .cl-prog{height:10px;background:#f1f1f1;border-radius:6px;overflow:hidden;margin:.4rem 0 .8rem}
  .cl-bar{height:100%;width:0%;background:#2b8cbe;transition:width .25s ease}
  .cl-stats{font-size:.9rem;color:#555}
</style>

<div class="cl-wrap" id="m2-checklist">
  <div class="cl-card">
    <h4>Module 2 — Checklist</h4>
    <div class="cl-note">
      Once you have watched the lecture recordings, completed the tutorial activity,
      and textbook problems, tick the outcomes you have achieved.
      Your selections are saved in your browser on this device.
    </div>

    <div class="cl-prog"><div class="cl-bar"></div></div>
    <div class="cl-stats"></div>

    <div class="cl-controls">
      <button class="btn" data-action="all">Mark all done</button>
      <button class="btn" data-action="clear">Clear all</button>
      <span class="cl-note">(* denotes expanded competencies)</span>
    </div>
  </div>

  <div class="cl-card">
    <div class="cl-group-title">Describing Distributions</div>

    <label class="cl-item">
      <input type="checkbox" data-id="d1">
      <span>Visually describe the distribution of a quantitative variable in terms of shape, centre, and spread from a histogram.</span>
    </label>

    <label class="cl-item">
      <input type="checkbox" data-id="d2">
      <span>Calculate by hand and interpret the median.</span>
    </label>

    <label class="cl-item">
      <input type="checkbox" data-id="d3">
      <span>Know the numbers that make up the five-number summary.</span>
    </label>

    <label class="cl-item">
      <input type="checkbox" data-id="d4">
      <span>Use technology to find the mean and standard deviation of a data set.</span>
    </label>

    <label class="cl-item">
      <input type="checkbox" data-id="d5">
      <span>Compare groups using boxplots.</span>
    </label>

    <label class="cl-item">
      <input type="checkbox" data-id="d6">
      <span class="cl-adv">* Calculate by hand the interquartile range (IQR) and five-number summary.</span>
    </label>

    <label class="cl-item">
      <input type="checkbox" data-id="d7">
      <span class="cl-adv">* Recognise the impact of outliers and skewness on measures of centre and spread.</span>
    </label>
  </div>
</div>

<script>
(function(){
  const ROOT = document.getElementById("m2-checklist");
  const KEY  = "m2-checklist-v1";

  const boxes = Array.from(ROOT.querySelectorAll("input[type=checkbox][data-id]"));
  const bar   = ROOT.querySelector(".cl-bar");
  const stats = ROOT.querySelector(".cl-stats");
  const btnAll   = ROOT.querySelector('[data-action="all"]');
  const btnClear = ROOT.querySelector('[data-action="clear"]');

  function update(){
    const total = boxes.length;
    const done  = boxes.filter(b => b.checked).length;
    const pct   = total ? Math.round(100 * done / total) : 0;
    bar.style.width = pct + "%";
    stats.textContent = `${done} of ${total} completed (${pct}%)`;
  }

  function save(){
    const state = {};
    boxes.forEach(b => state[b.dataset.id] = b.checked);
    localStorage.setItem(KEY, JSON.stringify(state));
    update();
  }

  function load(){
    const saved = JSON.parse(localStorage.getItem(KEY) || "{}");
    boxes.forEach(b => b.checked = !!saved[b.dataset.id]);
    update();
  }

  boxes.forEach(b => b.addEventListener("change", save));

  btnAll.addEventListener("click", () => {
    boxes.forEach(b => b.checked = true);
    save();
  });

  btnClear.addEventListener("click", () => {
    if (!confirm("Clear all ticks for this checklist on this device?")) return;
    localStorage.removeItem(KEY);
    boxes.forEach(b => b.checked = false);
    update();
  });

  load();
})();
</script>
```

## Using Statistical Software Checklist {-}

```{=html}
<div class="cl-wrap" id="m2-software-checklist">
  <div class="cl-card">
    <h4>Module 2 — Using Statistical Software Checklist</h4>
    <div class="cl-note">
      On successful completion of this module, students should be able to use SPSS or Excel
      to complete the following tasks.
    </div>

    <div class="cl-prog"><div class="cl-bar"></div></div>
    <div class="cl-stats"></div>

    <div class="cl-controls">
      <button class="btn" data-action="all">Mark all done</button>
      <button class="btn" data-action="clear">Clear all</button>
    </div>
  </div>

  <div class="cl-card">
    <div class="cl-group-title">Using Statistical Software</div>

    <label class="cl-item">
      <input type="checkbox" data-id="ss1">
      <span>Construct a histogram, a boxplot, and multiple boxplots.</span>
    </label>

    <label class="cl-item">
      <input type="checkbox" data-id="ss2">
      <span>Produce summary statistics of quantitative data.</span>
    </label>

    <label class="cl-item">
      <input type="checkbox" data-id="ss3">
      <span>Select subsets of data using Select Cases.</span>
    </label>
  </div>
</div>

<script>
(function(){
  const ROOT = document.getElementById("m2-software-checklist");
  const KEY  = "m2-software-checklist-v1";

  const boxes = Array.from(ROOT.querySelectorAll("input[type=checkbox][data-id]"));
  const bar   = ROOT.querySelector(".cl-bar");
  const stats = ROOT.querySelector(".cl-stats");
  const btnAll   = ROOT.querySelector('[data-action="all"]');
  const btnClear = ROOT.querySelector('[data-action="clear"]');

  function update(){
    const total = boxes.length;
    const done  = boxes.filter(b => b.checked).length;
    const pct   = total ? Math.round(100 * done / total) : 0;
    bar.style.width = pct + "%";
    stats.textContent = `${done} of ${total} completed (${pct}%)`;
  }

  function save(){
    const state = {};
    boxes.forEach(b => state[b.dataset.id] = b.checked);
    localStorage.setItem(KEY, JSON.stringify(state));
    update();
  }

  function load(){
    const saved = JSON.parse(localStorage.getItem(KEY) || "{}");
    boxes.forEach(b => b.checked = !!saved[b.dataset.id]);
    update();
  }

  boxes.forEach(b => b.addEventListener("change", save));

  btnAll.addEventListener("click", () => {
    boxes.forEach(b => b.checked = true);
    save();
  });

  btnClear.addEventListener("click", () => {
    if (!confirm("Clear all ticks for this checklist on this device?")) return;
    localStorage.removeItem(KEY);
    boxes.forEach(b => b.checked = false);
    update();
  });

  load();
})();
</script>
```
